Real-Time Certified Probabilistic Pedestrian Forecasting

### 2017-02-15 pedestrian_forecasting_iros_2017.tex
minor changes to implementation portion of text.

  \subsection{Learning the Vector Fields}
  Before we begin to learn vector-fields, we must learn the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
  In particular, we use the start and end point for each trajectory to obtain a point in $\mathbb{R}^4$.
  We then cluster in $\mathbb{R}^4$ using Affinity propagation \cite{FreyDueck2007}.
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

  For each set $S_k$ we learn a vector-field that is approximately compatible with that set.
  %We'd like to only consider vector-fields who's vectors have unit magnitude because we observe many of the trajectories in the data appear to have a roughly constant speed.
  Since most trajectories appearing in the dataset have roughly constant speed, we chose a vector-field that has unit magnitude. 
  That is, we assume the vector-field takes the form $X_k(x) = \left( \cos( \Theta_k(x) ) , \sin(\Theta_k(x)) \right)$ for some scalar function $\Theta_k(x)$.
  Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
  We assume $\Theta_k$ takes the form
  \begin{align*}
  	\Theta_k(x) = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x),
  \end{align*}
  for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
  In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
 $\Theta_k$ can be learned by computing the velocities observed in the cluster, $S_k$.
  These velocities may be obtained by a low order finite difference formula.
  Upon normalizing the velocities, we obtain a unit-length velocity vectors, $v_{i,k}$, anchored at each point, $x_{i,k}$, of $S_k$.
  We learn $\Theta_k$ by defining the cost-function:
  \begin{align*}
  	C[ \Theta_k] = \sum_i \langle v_{i,k} , ( \cos(\Theta_k( x_{i,k}) , \sin( \Theta_k( x_{i,k} ) ) \rangle
  \end{align*}
  which penalizes $\Theta_k$ for producing a misalignment with the observed velocities at the observed points of $S_k$.
  When $\Theta_{k}$ includes high order polynomials (e.g. beyond 5th order), we also include a regularization term to bias the cost towards smoother outputs.
  Using the $H^1$-norm times a fixed scalar suffices as a regularization term.

@@

  We begin by considering the nonlinear models first.
%  That is to say, we are concerned with $M = (k,s)$ for an integer $k$ and scalar $s$.
%  We'd like to compute $\Pr( x_0 \mid k,s)$ for each cluster $k=1,\dots,n$ and each speed $s \in \mathbb{R}$.
  We make the modeling assumption that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
  Additionally, we assume that $s$ and $k$ are independent.
  This means that we only need to learn $\Pr( x_0 \mid k)$, $\Pr(k)$, and $\Pr(s)$.

  We let $\Pr(k) = (n+1)^{-1}$ and $\Pr(s) \sim \mathcal{U}( [-s_{\max}, s_{\max} ] )$  where $s_{\max}>0$ is the largest observed speed in the dataset.  
  This implies that $\Pr(lin) = (n+1)^{-1}$ as well.
  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.

%  To learn $\Pr( x_0 \mid k)$, we assume that its logarithm may be approximated by a low dimensional subspace of functions on $\mathbb{R}^2$.
%  That is to say, 
  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is given by:
  \begin{align*}
  	V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)
  \end{align*}
  for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
  We chose our basis functions to be the collection of tensor products from the first six Legendre polynomials, normalized to the size of the domain.
  Then, one may fit the coefficients $c_{\alpha}$ to the data by using a log-likelihood criterion.
  The resulting (convex) optimization problem takes the form:
  \begin{align*}
  	\mathbf{c}^* = \inf_{ |\mathbf{c}| } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
  \end{align*}
  Where the norm on $\mathbf{c}$ is a sup-norm.
  We can also bias this optimization towards more regular functions by adding a penalty to the cost function.
    Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.

@@

 \subsection{Evaluating Performance}
In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
We implemented our model as well as the evaluation code in Python 2.6. Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.

>>>>

  \subsection{Learning the Vector Fields}
 We begin by identifying the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
  In particular, we use the start and end point for each trajectory to obtain a point in $\mathbb{R}^4$.
  We then cluster in $\mathbb{R}^4$ using Affinity propagation \cite{FreyDueck2007}.
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

  For each set $S_k$ we learn a vector-field that is approximately compatible with that set.
  %We'd like to only consider vector-fields who's vectors have unit magnitude because we observe many of the trajectories in the data appear to have a roughly constant speed.
  Since most trajectories appearing in the dataset have roughly constant speed, we chose a vector-field that has unit magnitude. 
  That is, we assume the vector-field takes the form $X_k(x) = \left( \cos( \Theta_k(x) ) , \sin(\Theta_k(x)) \right)$ for some scalar function $\Theta_k(x)$.
  Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
  We assume $\Theta_k$ takes the form:
  \begin{align*}
  	\Theta_k(x) = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x),
  \end{align*}
  for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
  In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
 $\Theta_k$ is learned by computing the velocities observed in the cluster, $S_k$.
  These velocities are obtained by a low order finite difference formula.
  Upon normalizing the velocities, we obtain a unit-length velocity vectors, $v_{i,k}$, anchored at each point, $x_{i,k}$, of $S_k$.
  We learn $\Theta_k$ by defining the cost-function:
  \begin{align*}
  	C[ \Theta_k] = \sum_i \langle v_{i,k} , ( \cos(\Theta_k( x_{i,k}) , \sin( \Theta_k( x_{i,k} ) ) \rangle
  \end{align*}
  which penalizes $\Theta_k$ for producing a misalignment with the observed velocities at the observed points of $S_k$.
  When $\Theta_{k}$ includes high order polynomials (e.g. beyond 5th order), we also include a regularization term to bias the cost towards smoother outputs.
  Using the $H^1$-norm times a fixed scalar suffices as a regularization term.

@@

We first considering the nonlinear models.
%  That is to say, we are concerned with $M = (k,s)$ for an integer $k$ and scalar $s$.
%  We'd like to compute $\Pr( x_0 \mid k,s)$ for each cluster $k=1,\dots,n$ and each speed $s \in \mathbb{R}$.
  We begin by assuming that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
  Additionally, we assume that $s$ and $k$ are independent.
  This means that we only need to learn $\Pr( x_0 \mid k)$, $\Pr(k)$, and $\Pr(s)$.

  We let $\Pr(k) = (n+1)^{-1}$ and $\Pr(s) \sim \mathcal{U}( [-s_{\max}, s_{\max} ] )$  where $s_{\max}>0$ is the largest observed speed in the dataset.  
  This implies that $\Pr(lin) = (n+1)^{-1}$ as well.
%  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.

%  To learn $\Pr( x_0 \mid k)$, we assume that its logarithm may be approximated by a low dimensional subspace of functions on $\mathbb{R}^2$.
%  That is to say, 
  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is given by:
  \begin{align*}
  	V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)
  \end{align*}
  for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
  We chose our basis functions to be the collection of tensor products of the first six Legendre polynomials, normalized to the size of the domain.
  Then, one may fit the coefficients $c_{\alpha}$ to the data by using a log-likelihood criterion.
  The resulting (convex) optimization problem takes the form:
  \begin{align*}
  	\mathbf{c}^* = \inf_{ |\mathbf{c}| } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
  \end{align*}
  Where the norm on $\mathbf{c}$ is a sup-norm.
  We bias this optimization towards more regular functions by adding a penalty to the cost function.
    Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.

@@

 \subsection{Evaluating Performance}
In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
We implement our model as well as the evaluation code in Python 2.6. 
Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.
